---
title: "Areal Data Model Fitting, Part 2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(maps)
library(maptools)
library(sf)
library(usmap)
library(arm)
library(CARBayes)
library(spdep)
```


Continuing with the Tester election dataset.

```{r, echo = F}
Tester <- read_csv('Tester_Results.csv')
Tester <- Tester %>% 
  mutate(Tester_Prop = TESTER / (TESTER + ROSENDALE + BRECKENRIDGE),
         county_fips = as.character(FIPS))

MT.counties <- map('county', 'montana', fill = T, plot = F)
#map('county', 'montana')
county.ID <- sapply(strsplit(MT.counties$names, ','), 
                    function(x) x[2])
mt.poly <- map2SpatialPolygons(MT.counties, IDs = county.ID)

#drop Yellowstone National Park
mt.poly.noYNP <- mt.poly[1:56,]
mt.nb.noYNP <- poly2nb(mt.poly.noYNP)
mt.listw <- nb2listw(mt.nb.noYNP, style = 'B')
mt.adj.mat <- nb2mat(mt.nb.noYNP, style = 'B')
```

## Now consider some covariates to explain the response

Consider a linear model with county population, using both `lm` and `S.glm`

```{r, echo = T}
lm_dat <- Tester %>% 
  left_join(usmap::countypop %>% rename("county_fips" = fips), by = 'county_fips') %>%
  mutate(scale_pop = scale(pop_2015))

pop_model <- lm(Tester_Prop ~  scale_pop, data = lm_dat)
display(pop_model)

S.glm(Tester_Prop ~ scale_pop, data = lm_dat, family = 'gaussian', 
      burnin = 100, n.sample = 1000, verbose = F)
```

\newpage

We previously extract the residuals create a choropleth and test for spatial association. Now we will directly run the CAR model using `CARBayes`. See [vignette](https://cran.r-project.org/web/packages/CARBayes/vignettes/CARBayes.pdf) for more info and options on model fitting, including GLM models.

This model can be expressed as

$Y_k \sim N(\mu_k, \nu^2),$ where $\mu_k = x_k \beta + \psi_k$.

A CAR prior is placed on $\psi_k$ such that the full conditional is expressed as

$$\psi_k|- \sim N \left( \frac{\rho \sum_i w_{ki}\psi_i}{\rho \sum_i w_{ki} + 1 - \rho}, \frac{\tau^2}{\rho \sum_i w_{ki} + 1 - \rho}\right)$$

If $\rho = 0$ there is no spatial structure present, if $\rho \rightarrow 1$ this is the intrinsic CAR model.



```{r}

S.CARleroux(Tester_Prop ~ scale_pop, data = lm_dat, family = 'gaussian', burnin = 10000, 
            n.sample = 100000, W = mt.adj.mat, thin = 5, verbose = F)
```

